\addchap{Acknowledgements}

I would like to thank my supervisor, Stuart Parkin, for
giving me an opportunity to work with him. I also appreciate his
encouragement, continuous guidance, laboratory facilities and help in different ways
throughout my stay in Halle.
\\
I would like to express my gratitude towards Amilcar Bedoya Pinto for continuous
counselling and guidance in planning my experiments. I would also like thank Kai Chang
and Ilya Kostanovskiy for their help in setting up the PAPAYA system in the begining
as well as teaching many experimental techniques. I had many enriching conversations 
with Frank Schuman which taught me a great deal about life, science and life in science.
\\
For technical assistance, I heartily acknowledge Kai-Uwe Assmann, Kornelia Sklarek, 
Martin Gottschalk, Norbet Schammelt, Mike Borrmann and Pierre-Jean Zermatten, 
as well as the wisdom of Winfried Heichler and Herbert Engelhard.
\\
I would like to extend my gratitude to Simone Jaeger and Antje Paetzold for helping
in administrative and personal matters without which I would not have had the conducive atmosphere
to complete my research. Simone also deserves special thanks for helping edit ths thesis.
\\
Here, my gratitude goes to the many 
other scientists and students for their advice and encouragement in
my development as a scientist: Fasil Kidane Dejene, Kumari Gaurav Rana, Neeraj Kumar, Peter 
Werner, Hakan Deniz, Holger Meyerheim, Robin Blaesing, Pranava Keerthi Sivakumar, Shouying 
Yang, James Taylor, Tianping Ma, Bharat Grover, Binoy 
Krishna Hazra, Banabir Pal, Hao Yang, Jue Huang, Yicheng Guan, Ankit Kumar Sharma, Abhay 
Kant Srivastava, Jibo Zhang, Jae-Chun Jeon, Andrea Migliorini, Kai-Uwe Demasius, Elena 
Derunova, Mazhar Ali, Amir Capua, Rana Saha, Fang Gao, Ajaya Nayak, Jiho Yoon, 
Alessandro Fumarola and Amine Wahada.
\\
I wish to record my deep sense of appreciation to our loving parents, family members and friends
for their moral support and care.